---
title: "Reddit Challenge 3"
author: "Matt Demas"
date: "1/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Challenge #3 "Credit Card Basics"
 * XLS dataset here: https://archive.ics.uci.edu/ml/machine-learning-databases/00350/default%20of%20credit%20card%20clients.xls
 * Description of data here: https://archive.ics.uci.edu/ml/datasets/default+of+credit+card+clients

This dataset covers 30,000 Taiwanese people with credit card debt in 2005 and whether they defaulted. For this challenge, we want some basic summary statistics.

Provide the median amount of credit given(LIMIT_BAL) and the likelihood of default (# of defaults/# of credit cards issued) for each of the following groups:
 * Male vs Female
 * High school vs college and higher
 * Age 30-39 vs 45-55

**Bonus:** Make your code minimalist. The fewer lines and the fewer characters, the better.

# Import the Data from the UCI Repository
```{r libraries}
library(readxl)
library(dplyr)
library(stats)
```

```{r import}
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/00350/default%20of%20credit%20card%20clients.xls"
destfile <- "df.xls"
download.file(url, destfile)
df <- read_excel(destfile, skip = 1)
df <- rename(df, y = `default payment next month`)
```

## Male vs. Female
```{r male-vs-female, echo=FALSE}
df %>% group_by(SEX) %>% summarize(median = median(LIMIT_BAL), pr_def = mean(y))
```

## High School vs. College and Higher
I will compare the education column with entries labeled as 3 (for high school) against those labeled 1 or 2 (for college and graduate school).

```{r edu_grouping_function_and_test, echo=FALSE}
# Add column to indicate college or greater vs. no college/unspecified
classify_education <- function(edu) {
  sapply(edu, function (edu) {
    if (edu == 1 || edu == 2) {
      return('college')
    } else if (edu == 3) {
      return('high school')
    } else if (edu == 4) {
      return('other')
    } else {
      return(NA)
    }
  })
}

testx = c(1, 2, 3, 4, -1, 10)
resultx = c('college', 'college', 'high school', 'other', NA, NA)
data.frame(classify_education(testx), resultx, classify_education(testx) == resultx)
```

```{r male-vs-female, echo=FALSE}
df %>% 
  mutate(EDUclass = classify_education(EDUCATION)) %>%
  group_by(COLLEGE) %>%
  summarize(median = median(LIMIT_BAL), pr_def = mean(y))
```

## Ages 30-39 vs. 45-55
I will compare the education column with entries labeled as 3 (for high school) against those labeled 1 or 2 (for college and graduate school).
```{r age_grouping_function_and_test, echo=FALSE}
classify_age <- function(age) {
  sapply(age, function (age) {
    if (age >= 30 && age <= 39) {
      return('30-39')
    } else if (age >= 45 && age <= 55) {
      return('45-55')
    } else {
      return('other')
    }
  })
}

testx = c(1, 29, 30, 31, 38, 39, 40, 44, 45, 54, 55, 56)
resultx = c('other', 'other', '30-39', '30-39', '30-39', '30-39', 'other', 'other', '45-55', '45-55', '45-55', 'other')
classify_age(testx) == resultx
```

```{r age_grouping_result}
df %>% 
  mutate(AGEclass = classify_age(AGE)) %>%
  filter(AGEclass != 'other') %>%
  group_by(AGEclass) %>%
  summarize(median = median(LIMIT_BAL), pr_def = mean(y))
```

# Visualizing Relationships with a Mosaic Plot 
I interpret the Interpret Probability of Default as Count
```{r}
df <- df %>%
  mutate(EDUclass = classify_education(EDUCATION)) %>%
  mutate(AGEclass = classify_age(AGE))
df_tab <- xtabs(~y + SEX + EDUclass + AGEclass, data = df)
```

```{r mosaicplot}
mosaicplot(df_tab, color = TRUE, shade = TRUE)
```